A Neural Network Alternative to Convolutive Audio Models for Source Separation

### 2017-06-13 paper/paper.tex
Updated paper and pdf to latest version

\maketitle
%
\begin{abstract}
We present a convolutiional auto-encoder that can act as a neural network equivalent to convolutive audio models. We demonstrate the ability of the network to learn optimal bases that resemble phoneme magnitude-spectra and also show how it can be used for supervised separation of speech mixtures. In terms of separation performance, the ability of the network to learn cross-frame characteristics from input spectrograms enables these models to significantly outperform their feed-forward counterpart. More importantly, we can now exploit the available diversity of neural networks to propose novel and potentially powerful extensions to existing auto-encoder architectures, based on a cascade of recurrent and convolutional neural network layers. \\

@@

We see that the CAE models significantly out-perform their corresponding FF versions. This can be seen from the fact that the inter-quartile range in SDR for CAE models is higher than the inter-quartile range of corresponding FF models for several values of $K$. This improvement is a consequence of reduced interference generated by these models in source separation (as seen in the SIR plots). Although the SAR for CAE models degrades slightly as compared to FF models, the significant improvement in SIR compensates for this loss. The performance of the models achieves a peak value for $K=80$. However, the separation performance does not degrade significantly for other values of $K$. Thus, the choice of $K$ does not appear to be a very critical consideration for these models unlike FF models. \\

The convolutive speech models obtained by training the RCAE also outperform the FF auto-encoder models, as seen by the median values and inter-quartile ranges. This improvement in performance is not as pronounced as the CAE models for the case of speech mixtures. However, the experimentation serves as a proof of concept that exploring non-uniform auto-encoder architectures could lead to interesting and potentially powerful algorithms for other datasets and applications. \\

>>>>

\maketitle
%
\begin{abstract}
We present a convolutional auto-encoder that can act as a neural network equivalent to convolutive audio models. We demonstrate the ability of the network to learn optimal bases that resemble phoneme magnitude-spectra and also show how it can be used for supervised separation of speech mixtures. In terms of separation performance, the ability of the network to learn cross-frame characteristics from input spectrograms enables these models to significantly outperform their feed-forward counterpart. More importantly, we can now exploit the available diversity of neural networks to propose novel and potentially powerful extensions to existing auto-encoder architectures, based on a cascade of recurrent and convolutional neural network layers. \\

@@

We see that the CAE models significantly out-perform their corresponding FF versions. This can be seen from the fact that the inter-quartile range in SDR for CAE models is higher than the inter-quartile range of corresponding FF models for several values of $K$. This improvement is a consequence of reduced interference generated by these models in source separation (as seen in the SIR plots). Although the SAR for CAE models degrades slightly as compared to FF models, the significant improvement in SIR compensates for this loss. The convolutive speech models obtained by training the RCAE also outperform the FF auto-encoder models significantly, as seen by the median values and inter-quartile ranges. This improvement in performance is not as pronounced as the CAE models for the case of speech mixtures. However, the experimentation serves as a definite proof of concept that exploring non-uniform auto-encoder architectures could lead to interesting and potentially powerful algorithms for other datasets and applications. \\

The performance of the models achieves a peak value for $K=80$. However, the separation performance does not degrade significantly for other values of $K$. Thus, the choice of $K$ does not appear to be a very critical consideration for these models unlike FF models. At the same time, the variance in SDR seems to be dependent on $K$. We observe that the variance in SDR is considerably lower for higher values of $K$ ($K\geq50$) as seen by the separation results for both the convolutive models. In other words, auto-encoder based convolutive audio models produce superior results consistently for a high number of CNN filters. \\
